\documentclass[11pt]{article}
\usepackage{fullpage,amsmath}

% --- -----------------------------------------------------------------
% --- Document-specific definitions.
% --- -----------------------------------------------------------------
\newtheorem{definition}{Definition}

\newcommand{\concat}{{\,\|\,}}
\newcommand{\bits}{\{0,1\}}

% --- -----------------------------------------------------------------
% --- The document starts here.
% --- -----------------------------------------------------------------
\begin{document}
\sloppy

\noindent Rutgers University\\
CS440: Introduction to Artificial Intelligence, Spring 2017\\
Kostas Bekris\\

\begin{center}
\LARGE{\textbf{Homework 2: James Carroll and Joel Carrillo}}\\
\large{\textbf{\emph{Classical, Local, and Adversarial Search; Constraint Satisfaction and Logic}}}
\end{center}

\vspace{.1in}

\begin{enumerate}

\item Problem 1
\begin{enumerate}
\item Statistics for the program are posted alongside this .pdf as .txt files. The 20-variable .txt also contains a link to the program's github.
\item 20 Variables
\begin{enumerate}
\item Average flips over 100 trials: 651
\item Average time taken over 100 trials: 3.24529519ms
\end{enumerate}
\item 50 Variables
\begin{enumerate}
\item Average flips over 100 trials: 11155
\item Average time taken over 100 trials: 49.23680068ms
\end{enumerate}
\item 75 Variables
\begin{enumerate}
\item Average flips over 100 trials: 12582
\item Average time taken over 100 trials: 82.21522391ms
\end{enumerate}
\item 100 Variables
\begin{enumerate}
\item Average flips over 100 trials: 392280
\item Average time taken over 100 trials: 3962.0381258299985ms
\end{enumerate}
\item There are two notable 'jumps' made in these calculations. Between 20 and 50, and 75 to 100.
\item There is a clear relationship between the number of variables and clauses in 3-CNF SAT problems.
\begin{enumerate}
\item More specifically, the number of different variables to track in each clause increases the flips required to satisfy all clauses.
\item Because of this, it takes more time. While the difference between 50 and 75 is relatively 'negligible', with only time doubling, the difference between 75 and 100 is far more significant.
\item 20 to 50 also indicates a rapid ramp-up, suggesting an exponential relationship that stabilizes at around 50-75, before increasing very rapidly at 100.
\end{enumerate} 
\end{enumerate}

\item Problem 2
\begin{enumerate}
\item Node values:
\begin{enumerate}
\item MIN(H-O): H = 8, I = 5, J = 6, K = 6, L = 18, M = 22, N = 8, O = 17
\item MAX(D-G): D = 8, E = 6, F = 22, G = 17
\item MIN(B-C): B = 6, C = 17
\item MAX(A): A = 17
\end{enumerate}
\item PRUNED: 16, 15
\begin{enumerate}
\item $H(8, \alpha= -\infty,\beta = 8)$, $I(5, \alpha = 8 , \beta = 5)$, $J(6, \alpha= -\infty,\beta = 6)$, $K(6, \alpha = 6, \beta = 6)$, $L(18, \alpha = 6, \beta = +\infty)$, $M(22, \alpha = 6, \beta = +\infty)$, $N(8, \alpha = 6, \beta = +\infty)$, $O(17, \alpha = 6, \beta = +\infty)$
\item $D(8, \alpha = 8, \beta= + \infty)$, $E(6, \alpha = 6, \beta = 8)$, $F(22, \alpha = 6, \beta = +\infty)$, $G(17, \alpha = 6, \beta = +\infty)$
\item $B(6, \alpha = -\infty, \beta = 6)$, $C(17, \alpha = 6 , \beta = +\infty)$
\item $A(17, \alpha = 17, \beta = 6)$
\end{enumerate}
\item Both the MAX players in either variation of minimax choose the action 17.
\begin{enumerate}
\item In alpha-beta pruning, the benefit is present in that once an action is found to be smaller than some given alpha value in a node of the same level, there no longer needs to be any further exploration. Even if there is some value lower further along in the child nodes, it will certainly not be considered for selection. Ultimately, because of the persistence of alpha and beta values, the maximum agent will always select its highest values possible, while the minimum agent will select the lowest values possible.
\end{enumerate}
\item A
\item Due to how the max selection works, certain values on the bottommost nodes will never be selected. For example, the left branch of I, 5, will never be considered as it will never make it past the larger H values, 8 or 99. The 6 attached to J and K will never make it past E, as the lowest possible value for D is 8. From this, we can see that values 5, 6, 8, 18 will never reach the top node, A.
\begin{enumerate}
\item Possible values for each node:
\begin{enumerate}
\item H = 99/8, I = 5/16, J = 83/6, K = 6/15, L = 18, M = 22/28, N = 99/8, O = 90/17
\item D = 8/16/99, E = 6/15/83, F = 22/28, G = 17/90/99
\item B = 8/16/83/99, C = 22/28/90/99
\item A = 22/28/83/90/99
\end{enumerate}
\item Alpha-beta pruning is \textit{not} possible for this due to the fact that there is no certainty that MIN will play optimally - it certainly would not work for the original tree.
\end{enumerate}
\end{enumerate}

\item Problem 3
\begin{enumerate}
\item Definitions
\begin{enumerate}
\item The set of variables are the 81 numbers: $\{n_{1,1}, \ldots, n_{1,9}, \ldots, n_{9, 1}, \ldots, n_{9, 9}\}$
\item Domain of possible values for each is $\{1, 2, 3, 4, 5, 7, 8, 9\}$
\item Constraints:
\begin{enumerate}
\item You must use the numbers already assigned.
\item $Alldiff( n_{1,1}, \ldots, n_{1, 9} )$ etc for all rows
\item $Alldiff( n_{1,1}, \ldots, n_{9, 1} )$ etc for all columns
\item Each of the 3x3 boxes must contain all the digits (none repeated).
\end{enumerate}
\end{enumerate}
\item Incremental formulation
\begin{enumerate}
\item Definitions
\begin{enumerate}
\item Start state: The board with M numbers specified
\item Successor function: Fill an unfilled cell
\item Goal test: Are there no repeats of numbers within a row, col, or box? 
\item Path cost function: How many times are the constraints violated? 
\item Heuristic: Minimum remaining values (to be less likely to choose incorrectly)
\item Branching factor: 9
\item Solution depth: 81 - M
\item Maximum depth: 81 - M
\item State space size: $(81-M)!$
\end{enumerate}
\end{enumerate}
\item An "easy" Sudoku problem allows little freedom once you have placed a number in a box, enabling Most-Restricted-Value to quickly end the game. A "hard" Sudoku problem allows more freedom.
\item Steps for solving Sudoku using local search:
\begin{enumerate}
\item Fill each empty cell so that there are 9 of each digit in each 3x3 box.
\item While constraints are not met:
\item Swap two numbers (not fixed as part of M) within a box
\item Heuristic is number of constraints violated in each row and column
\item Keep the change if heuristic decreases; discard if it increases.
\end{enumerate}
\item It will work better than the best incremental search algorithm on easy problems, but worse on hard problems, because the heuristic in local search tolerates more gradual improvements than MRV in incremental search.
\end{enumerate}

\item Problem 4
\begin{enumerate}
\item $((A \land K) \to D) \land (L \to K) \land (L \to W) \land (W \to !A)$
\begin{enumerate}
\item D: Superman is defeated
\item A: Superman is facing an opponent alone
\item K: Superman's opponent is carrying Kryptonite
\item L: Batman coordinates with Lex Luther
\item W: Wonder Woman fights on the side of Superman
\end{enumerate}
\item $(!A \lor !K \lor D) \land (!L \lor K) \land (!L \lor W) \land (!W \lor !A)$
\begin{enumerate}
\item Given: $((A \land K) \to D) \land (L \to K) \land (L \to W) \land (W \to !A)$
\item Definition of $\to$: $(!(A \land K) \lor D) \land (!L \lor K) \land (!L \lor W) \land (!W \lor !A)$
\item DeMorgan's rule: $(!A \lor !K \lor D) \land (!L \lor K) \land (!L \lor W) \land (!W \lor !A)$
\end{enumerate}
\item Desired solution $!D$. Negate for proof by contradiction: so assume $D$.
\begin{enumerate}
\item $(D \lor !K)$ and $(K \lor !L)$ so $!L$. (Resolution rule)
\item $(!L \lor W)$ and $(!W \lor !A)$ so $!A$ (Resolution rule)
\item $!A$ and $((A \land K) \to D)$ so $D$ is false.
\end{enumerate}
\end{enumerate}

\item Problem 5
\begin{enumerate}
\item Simulated annealing is useful for programs where there are a lot of variables: in this case, it becomes more useful as the number of cities increase in the Traveling Salesman problem. While the time increases for A* search would increase exponentially, with simulated annealing:
\begin{enumerate}
\item 10 cities: each run stopped at 1.8, 2.6, 3.6, 4.4, 5.1, 6.2, 7.9, 8.8, 9.5, 10.2, 11.1, 12.0, 13.6, 14.4, 15.2, 16.0, 17.3, 18.1, 19.0, 19.8, 21.3, 24.2, 27.0 seconds respectively. About 1 second each.
\item 25 cities: took 43 seconds in total for 25 lists of 25 cities, so average 1.72 seconds.
\item 50 cities: took 1 minute 9 seconds, so 2.76 seconds per solution.
\item 100 cities: took 3 minutes 12 seconds, so 1.92 seconds per solution
\end{enumerate}
\end{enumerate}

\item Problem 6
\begin{enumerate}
\item If both $h_{1}(n)$ and $h_2(n)$ are admissible, then their min is admissible, since $h(n) < h*(n)$.
\item With two admissible heuristics, their max is also admissible. If they are both consistent, then their max is also consistent.
\item ?
\item ?
\end{enumerate}


\end{enumerate}
\end{document}
